Exploring dimensions of biodiversity in Japanese ferns

### 2021-04-16 ms/manuscript.Rmd
Update MS

Ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are very densely sampled. Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.

@@

### Trait data

@@

@@

Observed values of PD, FD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others. Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values. Statistical significance was assessed using a two-tailed test with alpha = 0.5; observed values > 97.5% or < 0.25% were considered significant.

@@

We compared patterns in biodiversity between bioregions by calculating the standard effect size (SES) of biodiversity metrics. SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):

We compared SES of PD, RPD, FD, and RFD across bioregions using **FIXME: add this when figure out spatial-corrected statistical method to use**.

@@

@@

Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan. Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics. These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions. Finally, we consider the current conservation status of ferns in Japan and the implications of our findings for future conservation activity.

>>>>

The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to effect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019]. Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.

@@

### Morphological trait data

@@

### Reproductive mode data

We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa.

@@

Observed values of PD, FD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others. Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values. Statistical significance was assessed using a two-tailed test with alpha = 0.5; observed values > 97.5% or < 0.25% were considered significant. We also calculated the standard effect size (SES) of each biodiversity metric. SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):

We tested for a relationship between SES of each biodiversity metric and reproductive mode (% apomictic taxa) initially with linear models constructed using the R function 'lm'. Next, we checked for evidence of spatial autocorrelation in the linear model residuals with Moran's I using 'moran.mc' function in the spdep package. Spatial autocorrelatoin was frequently detected, so to acccount for this we constructed spatial linear mixed models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014]. We then again checked for any remaining spatial autocorrelation in the mixed models with Moran's I.

@@

@@

### Correlations between biodiversity and reproductive mode

Spatial autocorrelation as measured with Moran's I was detected in linear models, but was not in any of the spatial mixed linear models. SES of RPD in particular shows a strong negative correlation with % apomictic taxa (_ estimated effect _, SD).

@@

Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan. Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics, and suggest reproductive mode as an important driver in some cases. These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions. Finally, we consider the current conservation status of ferns in Japan and the implications of our findings for future conservation activity.
